---
title: "Final Project"
author: "Chiemela Onuoha"
format: revealjs
editor: visual
execute:
  echo: false
---

## Introduction

- For this project, I am using a dataset consisting of listings of over 8,000 movies and TV shows available on the Netflix streaming platform.  
- I chose this data because enjoy inspecting catalogs of shows and movies, so this dataset was a great option for me. There is a bit of a challenge with the data as well, allowing me to touch up on my data wrangling skills.  
- The dataset has 12 variables and 8,807 observations, with recorded titles added from 2008 to 2021.


## Data Source

- Data was found on [Kaggle.com](https://www.kaggle.com/datasets/shivamb/netflix-shows)^[Accessed November 2nd, 2025] from user "shivamb". 

## Univariate Analysis {.scrollable}

::: {.content-center}
```{r}
library(tidyverse)
library(gt)
library(gtExtras)
netflix <- read_csv("data/netflix_titles.csv")

# distinct genres
netflix |>
  separate_rows(listed_in, sep = ",\\s*") |>
  distinct(listed_in) |>
  filter(!is.na(listed_in) & listed_in != "NA") |>
  gt() |>
   tab_header(
    title = md("**Distinct Genres for Listings**")
   ) |>
  tab_source_note(
    source_note = md("Source: *Kaggle.com*")
    ) |>
  tab_options(
    column_labels.hidden = TRUE   
  )

```
:::

## Univariate Analysis

```{r}
# number of times each genre/category appears in dataset 
genre_counts <- netflix |>
  separate_rows(listed_in, sep = ",\\s*") |>
  count(listed_in, sort = TRUE) |>
  na.omit() 

# plot of genre counts
genre_counts |>
  ggplot(aes(listed_in, n)) +
  geom_col(fill = "antiquewhite4") +
  labs(
    x = "Genres",
    y = "Number of Titles",
    title = "Distribution of Distinct Genre Counts",
    caption = "Data Sourced from Kaggle.com"
  ) +
  theme_classic() +
  theme(
    axis.text.x = element_text(
      angle = 45,
      hjust = 1,
      size = 8,
      face = "bold"
      ),
    plot.title = element_text(size = 18)
  )
```

## Univariate Analysis

```{r}

year_counts <- netflix |>
  mutate(
    date_added = lubridate::mdy(date_added), # converting col to date type
    year_added = lubridate::year(date_added) # extracting year added
    ) |> # counting frequency years appear in dataset
  count(year_added, sort = TRUE) |>
  na.omit()

year_counts |> # plotting frequencies
  ggplot(aes(year_added, n)) +
  geom_col(fill = "#4E2A84") +
  scale_x_continuous(
    breaks = seq(2008, 2021, 1),
    expand = c(0, 0)
  ) +
  labs(
    title = "Bar Chart Showing Counts of Titles Added Each Year (2008-2021)",
    caption = "Data Sourced from Kaggle.com",
    x = "Year Added",
    y = "Number of Titles"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 18)
  )
```

## Bivariate Analysis

```{r}
# type of listing vs year added ----
netflix |>
  mutate(
    date_added = lubridate::mdy(date_added), # converting col to date type
    year_added = lubridate::year(date_added) # extracting year added
  ) |>
  na.omit() |>
  count(type, year_added) |>
  ggplot(aes(year_added, n, color = type)) +
  geom_line() +
  geom_point(aes(shape = type)) +
  scale_x_continuous(
    breaks = seq(2008, 2021, 1), # range for plot
    expand = c(0.01, 0) # extra space
    ) +
  scale_color_manual(
    values = c("Movie" = "#FF9F9B", "TV Show" = "#A1C9F4"), # changing colors for lines
    labels = c("Movie", "TV Show"), # adding labels for the legend
    name = NULL # remove legend title
  ) +
  scale_shape_manual(
    values = c("Movie" = 16, "TV Show" = 17), # changing shapes for lines
    labels = c("Movie", "TV Show"), 
    name = NULL
  ) +
  labs(
    title = "Movies vs TV Shows Added to Netflix Over Time",
    caption = "Data Sourced form Kaggle.com",
    x = "Year Added",
    y = "Number of Titles"
  ) +
  guides(color = guide_legend(
    override.aes = list(size = 2.5, stroke = 0.5))) +
  theme_classic() +
  theme(
    legend.position = c(0.5, 1),
    legend.justification = c(0.5, 1), # legend positioning
    legend.direction = "horizontal",
    legend.text = element_text(size = 11.5),
    plot.title = element_text(size = 18)
  ) 

```

## Bivariate Analysis

```{r}
netflix |>
  ggplot(aes(x = rating, fill = type)) +
  geom_bar(position = "dodge") +
  scale_fill_manual(
    values = c("Movie" = "#B6ACD1", "TV Show" = "#4E2A84"),
    labels = c("Movie", "TV Show")
  ) +
  labs(
    title = "How Netflix Ratings Differ Between Movies and TV Shows",
    caption = "Data Sourced from Kaggle.com",
    x = "Rating",
    y = "Number of Titles",
    fill = "Type of Listing"
    ) + 
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    plot.title = element_text(size = 18)
  )

```

