---
title: "Visualizing Netflix Data"
author: "Chiemela Onuoha"
format: revealjs
editor: visual
execute:
  echo: false
---

## Introduction

-   For this project, I am using a dataset consisting of listings of over 8,000 movies and TV shows available on the Netflix streaming platform.\
-   I chose this data because enjoy inspecting catalogs of shows and movies, so this dataset was a great option for me. There is a bit of a challenge with the data as well, allowing me to touch up on my data wrangling skills.\
-   The dataset has 12 variables and 8,807 observations, with recorded titles added from 2008 to 2021.

## Github Repo Link
[Chiemela Onuoha's Github](https://github.com/chiemelaonu/final-project-chiemelaonu)

## Data Source

-   Data was found on [Kaggle.com](https://www.kaggle.com/datasets/shivamb/netflix-shows)[^1] from user "shivamb".

[^1]: Accessed November 2nd, 2025

## Univariate Analysis {.scrollable .smaller}

There are 42 distinct genres in this dataset.

::: content-center
```{r}
#| label: tbl-1

# load packages
library(tidyverse)
library(gt)
library(gtExtras)
netflix <- read_csv("data/netflix_titles.csv")

# distinct genres
netflix |>
  separate_rows(listed_in, sep = ",\\s*") |>
  distinct(listed_in) |>
  filter(!is.na(listed_in) & listed_in != "NA") |>
  gt() |>
   tab_header(
    title = md("**Distinct Genres for Listings**")
   ) |>
  tab_source_note(
    source_note = md("Source: *Kaggle.com*")
    ) |>
  tab_options(
    column_labels.hidden = TRUE   
  )

```
:::

## Univariate Analysis {.smaller}

```{r}
#| label: plot-1

# number of times each genre/category appears in dataset 
genre_counts <- netflix |>
  separate_rows(listed_in, sep = ",\\s*") |>
  count(listed_in, sort = TRUE) |>
  na.omit() 

# plot of genre counts
genre_counts |>
  ggplot(aes(listed_in, n)) +
  geom_col(fill = "antiquewhite4") +
  labs(
    x = "Genres",
    y = "Number of Titles",
    title = "Distribution of Distinct Genre Counts",
    caption = "Data Sourced from Kaggle.com"
  ) +
  theme_classic() +
  theme(
    axis.text.x = element_text(
      angle = 45,
      hjust = 1,
      size = 8,
      face = "bold"
      ),
    plot.title = element_text(size = 18)
  )
```

Comedies, Dramas, and Independent Movies are the most common genres apparent in this dataset. Netflix appears to prioritize these genres possibly based on demand or from shared data that shows the 3 aforementioned genres as the more popular genres for movies.

## Univariate Analysis {.smaller}

```{r}
#| label: plot-2

year_counts <- netflix |>
  mutate(
    date_added = lubridate::mdy(date_added), # converting col to date type
    year_added = lubridate::year(date_added) # extracting year added
    ) |> # counting frequency years appear in dataset
  count(year_added, sort = TRUE) |>
  na.omit()

year_counts |> # plotting frequencies
  ggplot(aes(year_added, n)) +
  geom_col(fill = "#4E2A84") +
  scale_x_continuous(
    breaks = seq(2008, 2021, 1),
    expand = c(0, 0)
  ) +
  labs(
    title = "Bar Chart Showing Counts of Titles Added Each Year (2008-2021)",
    caption = "Data Sourced from Kaggle.com",
    x = "Year Added",
    y = "Number of Titles"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 18)
  )
```

The data records that most titles were added to Netflix between 2016 and 2021, with the highest amount being in 2019 with 2016 titles added.

## Bivariate Analysis {.smaller}

```{r}
#| label: plot-3

# type of listing vs year added ----
netflix |>
  mutate(
    date_added = lubridate::mdy(date_added), # converting col to date type
    year_added = lubridate::year(date_added) # extracting year added
  ) |>
  na.omit() |>
  count(type, year_added) |>
  ggplot(aes(year_added, n, color = type)) +
  geom_line() +
  geom_point(aes(shape = type)) +
  scale_x_continuous(
    breaks = seq(2008, 2021, 1), # range for plot
    expand = c(0.01, 0) # extra space
    ) +
  scale_color_manual(
    values = c("Movie" = "#FF9F9B", "TV Show" = "#A1C9F4"), # changing colors for lines
    labels = c("Movie", "TV Show"), # adding labels for the legend
    name = NULL # remove legend title
  ) +
  scale_shape_manual(
    values = c("Movie" = 16, "TV Show" = 17), # changing shapes for lines
    labels = c("Movie", "TV Show"), 
    name = NULL
  ) +
  labs(
    title = "Movies vs TV Shows Added to Netflix Over Time",
    caption = "Data Sourced form Kaggle.com",
    x = "Year Added",
    y = "Number of Titles"
  ) +
  guides(color = guide_legend(
    override.aes = list(size = 2.5, stroke = 0.5))) +
  theme_classic() +
  theme(
    legend.position = c(0.5, 1),
    legend.justification = c(0.5, 1), # legend positioning
    legend.direction = "horizontal",
    legend.text = element_text(size = 11.5),
    plot.title = element_text(size = 18)
  ) 

```

This plot shows the trends of how many Movie titles were added per year versus how many TV Show titles were added per year. Movies were added to Netflix at a higher frequency starting in 2014 and has increased drastically since then. Netflix focuses more on adding Movies to its service compared to TV Shows, which could be attributed to user demand or from trends of the most popular forms of media over time. More investigation with outside data/sources would be needed to confirm the reasons behind this trend.

## Bivariate Analysis {.smaller}

```{r}
#| label: plot-4

netflix |>
  ggplot(aes(x = rating, fill = type)) +
  geom_bar(position = "dodge") +
  scale_fill_manual(
    values = c("Movie" = "#B6ACD1", "TV Show" = "#4E2A84"),
    labels = c("Movie", "TV Show")
  ) +
  labs(
    title = "How Netflix Ratings Differ Between Movies and TV Shows",
    caption = "Data Sourced from Kaggle.com",
    x = "Rating",
    y = "Number of Titles",
    fill = "Type of Listing"
    ) + 
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    plot.title = element_text(size = 18)
  )

```

This plot shows how Netflix ratings differ between Movies and TV Shows on the platform. For both movies and TV shows, most of the listings have a rating of either TV-14 or TV-MA. Netflix's catalog targets a teen or young adult to adult audience when adding listings to the service.

## Conclusion {.smaller}

This project explored patterns in Netflix’s catalog by focusing on questions such as which genres appear most often, how the number of added titles has changed over time, and how movies and TV shows differ in frequency. The analysis showed that Comedy, Drama, and Independent Movies occur the most. The number of titles added rose sharply between 2016 and 2021, with 2019 having the highest count. Movies appear more often than TV shows throughout the dataset.

These findings give an initial picture of how Netflix may curate its catalog, with emphasis on popular genres and steady growth in content additions. However, this dataset represents only a limited view of the platform. More complete data and deeper research would be required to make stronger conclusions about Netflix’s long-term content strategy.
